### 2023-05-19 src/tex/ms.tex
[showyourwork] overleaf sync

\RequirePackage{cmap}

\documentclass[usenatbib]{mnras}

\usepackage[T1]{fontenc}
\usepackage[dvipsnames, x11names, svgnames]{xcolor}

\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}

\usepackage{xurl}
\urlstyle{tt}

\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[
    % colorlinks=true, 
    % allcolors=blue,
]{hyperref}

% Prettier fonts 
% Those seem to be the "right" fonts for MNRAS
\usepackage{newtxtext}
\usepackage{newtxmath}

\usepackage{microtype}
\usepackage{csquotes}
\usepackage[capitalise]{cleveref}
\usepackage{xspace}

\usepackage{standalone}
\usepackage{tikz}
\usetikzlibrary{
    calc,
    positioning,
    shapes,
    shapes.symbols,
}

\usepackage{siunitx}
\sisetup{
    range-units = single,
    range-phrase = \textup{--},
    per-mode = symbol,
}
\DeclareSIUnit\parsec{pc}
\DeclareSIUnit\au{au}
\DeclareSIUnit\mas{mas}

% Load journal macros (e.g., \aa for A&A)
% Source: https://ui.adsabs.harvard.edu/help/actions/journal-macros
\usepackage{aas_macros}
% \usepackage{astro_bib_macro}  % where is this one from?

\newcommand{\todo}[1]{\textcolor{red}{[#1]}}
\newcommand{\timmy}[1]{\textcolor{red}{[\textbf{Timmy:} #1]}} % amazing

% Define macros
\newcommand{\IWA}{\ensuremath{\mathrm{IWA}}\xspace}
\newcommand{\OWA}{\ensuremath{\mathrm{OWA}}\xspace}
\newcommand{\HWO}{HabWorlds\xspace}

% Define the colors of the \rainbows
\definecolor{C0}{HTML}{ff0000}
\definecolor{C1}{HTML}{ff6d38}
\definecolor{C2}{HTML}{ecc86f}
\definecolor{C3}{HTML}{a4f89f}
\definecolor{C4}{HTML}{5af8c8}
\definecolor{C5}{HTML}{12c8e6}
\definecolor{C6}{HTML}{386df9}
\definecolor{C7}{HTML}{8000ff}

\newcommand{\rainbows}{% MAK: KEEP this formatting for MNRAS submission. It deserves this!
    \textcolor{C0}{r}%
    \textcolor{C1}{a}%
    \textcolor{C2}{i}%
    \textcolor{C3}{n}%
    \textcolor{C4}{b}%
    \textcolor{C5}{o}%
    \textcolor{C6}{w}%
    \textcolor{C7}{s}%
    \xspace
}

\newcommand{\istar}{%
  \hspace{-0.125em}%
  \tikz[baseline=(char.base)]{
    \node [inner sep=0] (char) {i};
    \node[starburst, starburst points=5, starburst point height=2.5mm, fill=White, draw=MidnightBlue, line width=0.1pt, minimum width=0.3mm, minimum height=0.1mm, inner sep=0.9pt, rotate=15] at (0.0015, 0.15) {};
  }%
  \hspace{-0.175em}%
}
\newcommand{\oceanglints}{\textcolor{MidnightBlue}{ocean gl\istar{}nts}\xspace}

% Patch \hypertarget; for details, see 
% https://tex.stackexchange.com/a/17138/97669
\makeatletter
\newcommand{\affdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}$^\mathrm{#1}$}
\makeatother
\newcommand{\afflink}[1]{\hyperlink{#1}{#1}}

% Full title of the paper (MNRAS has no running title?)
\title{Chasing \rainbows and \oceanglints:\\ Inner working angle constraints for the Habitable Worlds Observatory}

% Authors are in alphabetical order:
% DO NOT EDIT THIS MANUALLY
% ALL ENTRIES ARE GENERATED BY A SCRIPT
% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
\author[Sophia R. Vaughan et al.]{%
    Sophia R. Vaughan\thanks{Correspondence:~\url{sophia.vaughan@physics.ox.ac.uk}}\textsuperscript{,\afflink{1}},
    Kimberly Bott\textsuperscript{\afflink{2},\afflink{3},\afflink{4}},
    Sarah L. Casewell\textsuperscript{\afflink{5}},
    Nicolas B. Cowan\textsuperscript{\afflink{6}},
    David S. Doelman\textsuperscript{\afflink{7},\afflink{8}},
    \newauthor
    Timothy D. Gebhard\textsuperscript{\afflink{9},\afflink{10}},
    Matthew Kenworthy\textsuperscript{\afflink{7}},
    Johan Mazoyer\textsuperscript{\afflink{11}},
    Maxwell A. Millar-Blanchaer\textsuperscript{\afflink{12}},
    \newauthor
    Olivier Absil\textsuperscript{\afflink{13}},
    Lisa Altinier\textsuperscript{\afflink{14}},
    Pierre Baudoz\textsuperscript{\afflink{11}},
    Ruslan Belikov\textsuperscript{\afflink{15}},
    Alexis Bidot\textsuperscript{\afflink{16}},
    Jayne Birkby\textsuperscript{\afflink{1}},
    \newauthor
    Markus J. Bonse\textsuperscript{\afflink{10}},
    Bernhard Brandl\textsuperscript{\afflink{7}},
    Alexis Carlotti\textsuperscript{\afflink{16}},
    Elodie Choquet\textsuperscript{\afflink{14}},
    Niyati Desai\textsuperscript{\afflink{17}},
    \newauthor
    Kevin Fogarty\textsuperscript{\afflink{15}},
    Jules Fowler\textsuperscript{\afflink{18}},
    Yann Gutierrez\textsuperscript{\afflink{11},\afflink{19},\afflink{20}},
    Olivier Guyon\textsuperscript{\afflink{21},\afflink{22},\afflink{23},\afflink{24}},
    Sebastiaan Y. Haffert\textsuperscript{\afflink{21}},
    \newauthor
    Olivier Herscovici-Schiller\textsuperscript{\afflink{19}},
    Adrien Hours\textsuperscript{\afflink{16}},
    Roser Juanola-Parramon\textsuperscript{\afflink{25},\afflink{26}},
    Elina Kleisioti\textsuperscript{\afflink{7},\afflink{27}},
    \newauthor
    Lorenzo König\textsuperscript{\afflink{13}},
    Mariya Krasteva\textsuperscript{\afflink{28}},
    Iva Laginja\textsuperscript{\afflink{11}},
    Rico Landman\textsuperscript{\afflink{7}},
    Lucie Leboulleux\textsuperscript{\afflink{16}},
    \newauthor
    David Mouillet\textsuperscript{\afflink{16}},
    Mamadou N’Diaye\textsuperscript{\afflink{29}},
    Emiel H. Por\textsuperscript{\afflink{30}},
    Laurent Pueyo\textsuperscript{\afflink{30}},
    Frans Snik\textsuperscript{\afflink{7}},
    \newauthor
    Daphne M. Stam\textsuperscript{\afflink{31}},
    Victor Trees\textsuperscript{\afflink{32},\afflink{33}},
    Dirk van Dam\textsuperscript{\afflink{7}},
    Kyle van Gorkom\textsuperscript{\afflink{21}},
    Maaike van Kooten\textsuperscript{\afflink{34}}
    \newauthor \\%
    Affiliations are listed at the end of the paper
    % Timmy: I checked MNRAS and it seems that this is how they handle very long author lists. See, for example:
    % https://academic.oup.com/mnras/article/498/2/2354/5900562
}
\date{Draft version: \today}

\pagestyle{plain}

\begin{document}

\maketitle

\begin{abstract}
NASA recently announced the Habitable Worlds Observatory (\HWO), a coronagraphic space mission to detect rocky planets in habitable zones, establish their habitability, and search them for biosignatures. 
%
Surface liquid water is central to the definition of planetary habitability.
%
Photometric and polarimetric phase curves of starlight reflected by an exoplanet can reveal ocean glint, rainbows and other phenomena caused by scattering by clouds or atmospheric gas.
%
Direct imaging missions are optimised for planets near quadrature, and \HWO' coronagraph may obscure the phase angles where such optical features are strongest. 
%
The range of accessible phase angles for a given exoplanet will depend on the planet's orbital inclination and/or the coronagraph's inner working angle (IWA). 
%
For planets in near edge-on orbits, accessible phase angles are limited by coronagraphic obscuration. 
%
We use \HWO'~catalogue of 160 target stars to estimate the number of exo-Earths that could be searched for ocean glint, rainbows, and Rayleigh scattering. 
%
We find that any exo-Earths in planetary systems in this catalogue will have an accessible Rayleigh scattering peak in their polarimetric phase curve. 
%
The rainbow due to water clouds at phase angles of $\sim\qtyrange{20}{60}{\degree}$ would be accessible to \HWO in $\sim\num{42}$ systems, while the ocean glint signature at phase angles of $\sim\qtyrange{130}{170}{\degree}$ would be accessible in $\sim\num{14}$ systems, assuming an \IWA of \qty{62}{\mas} ($3\,\lambda/D$ for a $D=\qty{6}{\meter}$ telescope at $\lambda=\qty{600}{\nano\meter}$).
%
Improving the \IWA to $2\,\lambda/D$ increases the number of systems with accessible rainbows and glints by factors of roughly 2 and 3, respectively.
%
The number of systems in which an exo-Earth could be searched for these scattering phenomena increases slightly for eccentric orbits.  %and scales approximately inversely with the \IWA.  
\end{abstract}

\begin{keywords}
    planets and satellites: terrestrial planets -- 
    instrumentation: high angular resolution -- 
    planets and satellites: atmospheres
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Liquid surface water is closely related to planetary habitability because it is essential for life as we know it.
%
To first order, its presence can be predicted by the flux a planet receives from its star.
%
The habitable zone (HZ) is defined as the range of semi-major axes around a given main sequence star, for which a rocky planet may have the right temperature to maintain liquid surface water \citep{kasting93}. 
%
However, there are many factors besides the incident stellar flux that influence a planet's ability to hold liquid surface water, such as the atmospheric pressure and geological activity. 
%
Detecting liquid surface water on an exoplanet would thus provide important context for the interpretation of possible biosignatures.
%
Variations of the flux and degree of polarization as an exoplanet orbits its host star---phase variations---could reveal scattering phenomena from condensed water \citep[e.g., rainbows;][]{bailey2007} or specular reflection, i.e. glint, from water oceans \citep[][]{2008Icar..195..927W}.
%
However, depending on a planetary system's architecture, such features can be obscured by the inner working angle (\IWA) of a telescope's coronagraph, shown schematically in Fig.~\ref{fig:annotated-orbit}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Angle conventions}

As shown in Fig.~\ref{fig:annotated-orbit}, we define the \emph{scattering angle} $\varphi$ as the planet--star--observer angle (large angles for back-scattering, small angles for forward-scattering) and the supplementary \emph{phase angle} $\alpha$ (zero at full phase and 180$^\circ$ at new phase). 
%
A planet on a perfectly face-on orbit is always seen at quadrature, $\varphi = \alpha = 90^\circ$, while a planet on an inclined orbit will exhibit a wider range of viewing angles; only perfectly edge-on orbits pass through all possible scattering angles. 
%
In some plots, we report \emph{degrees from quadrature}, so that both new and full phase are at \qty{90}{\degree} for edge-on orbits.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includestandalone{tikz/annotated-orbit}
    \caption{
        Illustration of a planet in an edge-on, circular orbit showing how a coronagraph can obscure the exoplanet at small and large phase angles, and thus prevent observations of characteristic scattering features in an exoplanet's phase curve. 
        The scattering angle (the planet--star--observer angle), $\varphi$, is shown in red, and the supplementary phase angle, $\alpha$, is shown in blue. 
        See also \cref{fig:scattering-angle} and \cref{tab:phase_ranges}.
    }
    \label{fig:annotated-orbit}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ocean glint}

Ocean glint is the specular (Fresnel) reflection of light off a smooth liquid surface. 
%
While glint is a familiar phenomenon on Earth, there is only one other body in the Solar System where it has been observed: the Cassini orbiter captured the glint of sunlight reflecting off the liquid hydrocarbon lakes near Titan's North Pole as it was observing Titan at crescent phases \citep{2010GeoRL..37.7104S}.
%
In the total flux, the glint signal is indeed expected to be most prominent at large planetary phase angles.
%
The maximum degree of polarization of the glint occurs at the Brewster angle, which is determined by the refractive indices at the interface; between air and liquid water this occurs at a reflection angle of \qty{53}{\degree} locally on the surface \citep[at a planetary phase angle of about \qty{106}%\qty{150}
{\degree}; see, e.g.,][]{Zugger_2010,treesandstam2019}.%{2008Icar..195..927W,Zugger_2010,treesandstam2019}
%
This angle therefore varies slightly with the wavelength and the composition of the ocean and the atmosphere.
%
The angular width of the glint in total and polarized flux generally broadens with increasing wave height \citep{2008Icar..195..927W, kopparla2018, treesandstam2019, trees2022}.\looseness=-1 \space In general, the glint on a planet will be strongest at red wavelengths because there Rayleigh scattering and attenuation in the atmosphere are smallest \citep{Zugger_2011}. 
%
This has been observed in Earthshine measurements \citep{Emde2017,sterzik2019, takahashi2021}. 
%
At large phase angles, the forward scattering of light by cloud particles could be confused with glint \citep{Robinson_2010}.
%
This confusion can be resolved by measuring polarized phase curves at multiple wavelengths \citep{treesandstam2019}, as shown in \cref{fig:bottplot}, or by probing 
the rainbow, see below.

Glint provides evidence of specular reflecting surfaces, which, taken in the context of the temperature and types of condensates, may be interpreted as evidence for liquid water.
%
The presence of liquid water on the surface of a planet implies the presence of a water cycle which provides key context to its habitability and the interpretation of atmospheric biosignatures.
%
Aside from detecting ocean glint, the presence of surface water on Earth-like exoplanets can be established through less direct methods near quadrature such as spectropolarimetry, to infer ocean reflection \citep{trees2022}, or through rotational mapping, to identify bright continents and dark oceans \citep[e.g.,][]{2009ApJ...700..915C,lustig2019}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics{figures/figure-1-bott-plot.pdf}
    \script{create-figure-1-bott-plot.py}
    \caption{
        The phase curves of the normalized linearly polarized flux of a cloudy planet with a wavy ocean surface and an Earth-like atmosphere in an edge-on orbit; adapted from \citet{treesandstam2019}.  
        %
        The wind speed over the ocean, which determines the wave height \citep[][]{CoxMunk1954}, is \qty{7}{\meter\per\second}.
        %
        The clouds consist of spherical water droplets.
        %
        The cloud coverage fraction is 0.5 and while the cloud pattern is patchy, the glint is always cloud-free. 
        %
        The fluxes are normalized such that the total (unpolarized + polarized) flux equals the planet's geometric albedo at phase angle $\alpha = 0^\circ$ (scattering angle $\varphi = 180^\circ$).
        %
        Below the plot, we show the scattering angles of glories, rainbows, Rayleigh scattering, and ocean glint. Note that the features in the actual curves partly overlap with each other.
    }
    \label{fig:bottplot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rainbows, cloudbows, and glories}

Rainbows, cloudbows, glories and other scattering phenomena can dramatically influence the shape of a planet's photometric and polarimetric phase curves and provide a wealth of information about exoplanets \citep{karalidi2012rainbow, stam2008,bailey2007,2014A&A...566L...1G}.%
\footnote{
    Rainbows occur when light is scattered by rain droplets that are large compared to the wavelength of the light. 
    Cloudbows refer to features occurring at the same phase angles from smaller cloud particles, with the spread in colours invisible to the human eye making them appear white \citep[for further discussion, see][]{bailey2007}.  
}
%
In the Solar System, the polarized back-scatter from Titan indicated that the moon's reflection was due to an atmospheric haze rather than a surface \citep{zellner1973polarization}, Venus's polarized phase curve at three wavelengths showed that the clouds were $\sim\qty{75}{\percent}$ H\textsubscript{2}SO\textsubscript{4}--water mixture and constrained the droplet size distribution \citep{hansenhovenier1974}.
%
The glory feature in Venus's flux phase curve was later used to rule out various species as the unknown UV absorber \citep{petrova2018glory}.
%
Glories and rainbows are technically visible both in total and polarized flux phase curves in the presence of spherical cloud and/or haze particles, but their strongly polarized nature coupled with the suppression of other polarized features at gibbous and full phases make them more easily characterized with polarimetry than with photometry \citep{karalidi2011, stam2008, treesandstam2019,trees2022}.
\looseness=-1
%

The phase angles and wavelength dependence of these features are sensitive to the chemical composition and droplet size distribution of condensates in the atmosphere. 
%
As with ocean glint, the detection of liquid condensates on an exoplanet can provide key context to their habitability, potential water cycle, and interpretation of atmospheric biosignatures.
%
In this paper, we define our optimal angle for the detection of a rainbow as that of the peak of the water droplet rainbow at visible wavelengths, which is around a phase angle $\alpha$ of \qty{40}{\degree} (see \cref{fig:bottplot}).
%
Pushing observations to angles closer to the star (smaller and larger phase angles) would allow the determination of the full shape of the rainbow, and enable the identification of rainbows due to the scattering of light by other condensates.
%
The glory can allow further characterization of condensates as its angle and shape also depend on the refractive index, but this glory occurs within a phase angle of \qty{20}{\degree}, when the on-sky separation is small, as can also be seen in \cref{fig:bottplot}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Habitable Worlds Observatory}

The National Academy of Sciences Astronomy \& Astrophysics 2020 Decadal Survey (Astro2020) \citep{decadal} recommended the first in a new \enquote{Great Observatories} program a telescope with the capability to detect signatures of habitability on about 25 HZ planets.
%
This requires an instrument with a coronagraph capable of high-contrast imaging and low-resolution spectroscopy at optical to near-infrared wavelengths. 
%
The precursor technology recommended by the survey lists \enquote{direct imaging to probe polarized ocean glint on terrestrial planets} as a priority capability \citep[Box E.1 in][]{decadal} for ground and space-based observatories.
%
Following the release of this report, NASA recently announced the start of the development of the Habitable Worlds Observatory (hereafter ``\HWO'').
%
The performance and characteristics (on- or off-axis, diameter, type of segmentation, polarimetry or photometry, and coronagraph, \IWA) of \HWO will significantly impact the expected exo-Earth yield \citep{Stark2019_exoplanetyield} and are still to be determined.
%
They will likely be heavily informed by the LUVOIR \citep{LUVOIR2019} and HabEx \citep{HabEx_2020} preparatory studies.
%
For generality, in this work, we report angular separations in milliarcseconds however, for ease of comparison, \cref{tab:IWA_OWA} lists the angular separations corresponding to multiples of $\lambda / D$ for a $\qty{6}{\meter}$ telescope, as suggested in Astro2020, at representative wavelengths $\lambda$ of $\qty{600}{\nano\meter}$ and $\qty{1}{\micro\meter}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
    \centering
    \caption{
        Inner and Outer Working Angles (\IWA and \OWA{}, respectively) in milli-arcseconds (\unit{\mas}) as functions of $\lambda / D$, at wavelengths $\lambda$ of \qty{600}{\nano\meter} and \qty{1}{\micro\meter}, assuming a telescope diameter $D$ of 6\,m.
    }
    \label{tab:IWA_OWA}
    \begin{tabular}{ c c c c } 
    \toprule
    & $\lambda/D$ & \unit{\mas} (at \qty{600}{\nano\meter}) & \unit{\mas} (at \qty{1}{\micro\meter}) \\
    \midrule
    \midrule
    IWA & 1 & 20.63 & 34.38 \\
    IWA & 2 & 41.25 & 68.75 \\
    IWA & 3 & 61.88 & 103.13 \\
    IWA & 4 & 82.51 & 137.51 \\
    \midrule
    OWA & 32 &  660.05 & 1100.08 \\
    OWA & 64 & 1320.09 & 2200.16 \\
    \bottomrule
    \end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Outline}

In this paper, we quantify the planetary phase angles that would be accessible for direct observations of hypothetical terrestrial planets orbiting in the habitable zones of \HWO target stars.
%
In \cref{subsec:2.1}, we describe the target list, in \cref{subsec:2.2}, we derive expressions for the range of phase angles given a few different limits, in \cref{subsec:2.3}, we describe Monte Carlo simulations to marginalize over the unknown orbital inclination and eccentricity of planets around these stars, and in \cref{subsec:2.4}, we outline how the photometric and polarimetric planet/star contrast changes along a planet's orbit.  
%
We present and discuss our results in \cref{sec:3}, and conclude in \cref{sec:4}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\HWO and its stellar sample}
\label{subsec:2.1}

A preliminary target list of nearby stars (maximum distance \qty{25}{\parsec}) around which \HWO may search for habitable rocky planets, assuming a \qty{6}{\meter} telescope, has been compiled: the \emph{NASA Exoplanet Exploration Program's Mission Star List for the Habitable Worlds Observatory}.%
\footnote{Available online: \url{https://exoplanets.nasa.gov/internal_resources/2645_NASA_ExEP_Target_List_HWO_Documentation_2023.pdf}}
%
This list consists of $\num{164}$ stars, the majority of which are Sun-like dwarfs: 66~F~dwarfs, 55~G~dwarfs, 40~K~dwarfs, and 3~M~dwarfs.
%
The target list is constructed assuming a maximum planet magnitude in the Cousins Rc band of $R_c = 31$ and a lower limit of \num{2.5e-11} for the planet-star flux ratio. 
% 
The adopted HZ limits are a semi-major axis of \qtyrange{0.95}{1.67}{\au} for a solar twin and planet sizes between \qtyrange{0.8}{1.4}{} Earth radii. 
%
For non-solar stars, the semi-major axis scales as square root of the bolometric luminosity normalized to the Sun. 
%
This range of orbital separation corresponds to the \enquote{conservative habitable zone} \citep{kasting93, kopparapu13}. 
%
In the remainder of this work, we consider hypothetical Earth-sized planets orbiting at the right distance from their parent star to receive the same flux as the Earth receives from the Sun.
%
By design, such planets should be observable by \HWO at quadrature phase.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Accessible phase angles}
\label{subsec:2.2}

We first define the phase angle ranges over which the various phenomena --- glint, Rayleigh peak, rainbow and glory --- manifest themselves in \cref{tab:phase_ranges}, also shown in Fig.~\ref{fig:bottplot}. 
%
These ranges are all approximate, as the exact angles where these phenomena dominate depend on the details of the atmospheric and surface properties, as well as on the wavelengths being observed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table 2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
    \centering
    \caption{
        The approximate ranges of the phase angle $\alpha$ for the different optical phenomena considered in this paper (cf. Fig.~\ref{fig:bottplot}).
    }
    \label{tab:phase_ranges}
    \begin{tabular}{ l c c c } 
        \toprule
        & $\alpha_\mathrm{min}$ & $\alpha_\mathrm{peak}$ & $\alpha_\mathrm{max}$ \\
        \midrule
        Glory       & \qty{0}{\degree}     & \qty{5}{\degree}     & \qty{10}{\degree} \\
        Rainbow    & \qty{22}{\degree}    & \qty{42}{\degree}    & \qty{63}{\degree} \\
        Rayleigh peak    & \qty{50}{\degree}    & \qty{70}{\degree}    & \qty{110}{\degree} \\
        Ocean Glint & \qty{130}{\degree}   & \qty{150}{\degree}   & \qty{170}{\degree} \\
        \bottomrule
    \end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A coronagraphic mask can obscure part of a planet's orbit. As illustrated in \cref{fig:annotated-orbit} obscuration is a problem at small and large phase angles, when the planet is at small angular separation from its star.
The range of observable phase angles will depend on the \IWA of the coronagraph, as well as on the inclination and eccentricity of the exoplanet's orbit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Circular, edge-on orbits}

For the non-eccentric case, the minimum ($\alpha_\mathrm{min}$) and maximum ($\alpha_\mathrm{max}$) accessible phase angles are symmetric about \qty{90}{\degree}. 
We define the angle $\Delta \alpha$ such that: 
\begin{equation}
    \label{eq:Delta_phi}
    \Delta \alpha 
    = \alpha_\mathrm{max} - \alpha_\mathrm{min}
    =  2 \cdot (\alpha_\mathrm{max} - \qty{90}{\degree}) 
    =  2 \cdot (\qty{90}{\degree} - \alpha_\mathrm{min} )  \,.
\end{equation} 
The on-sky projected planet distance, $r_\mathrm{proj}$ (in \unit{\au}) is shown in \cref{fig:scattering-angle} for a non-eccentric case with a semi-major axis $a$ (in \unit{\au}), at a distance $d_*$ (in \unit{\parsec}) from the observer.
%
The scattering angle, $\varphi$, is given by
\begin{equation}
    \label{eq:sinphi}
    \sin(\varphi) = - \sin(\alpha) = \frac{r_\mathrm{proj}}{a} \,,
\end{equation}
assuming small angles and $r_\mathrm{proj} = \delta d_*$, where $\delta$ is the on-sky angular separation.
%
We access the minimum phase angle when $r_\mathrm{proj}$ reaches the inner working angle separation, thus when $\delta = \mathrm{IWA}$. 
%
Combining \cref{eq:Delta_phi} and \cref{eq:sinphi}, we deduce:
\begin{equation}
    \label{eq:scattering_angle}
    \cos\left(\dfrac{\Delta \alpha}{2}\right) = \frac{\mathrm{IWA} 
    \cdot d_*}{a} \,.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 3
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includestandalone{tikz/scattering-angle}
    \caption{
        The angular separation of a planet from its star as seen by the observer, $\delta$, is related to the scattering angle, $\varphi$. The star is at a distance $d_*$ from the observer, the planet orbits a distance $a$ from the star, and  the projected distance between the star and planet is  $r_\mathrm{proj}$. Not to scale. 
    }
    \label{fig:scattering-angle}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Cref{fig:scatterplot} shows $\Delta \alpha$ for the systems in the \HWO target list, assuming an IWA of \qty{62}{\mas}.
%
As can be seen in \cref{fig:scatterplot}, for exoplanets in edge-on circular orbits, it will always be possible to observe the peak polarization due to Rayleigh scattering, which occurs around quadrature ($\alpha=\qty{90}{\degree}$), while the rainbow angle (around $\alpha=\qty{40}{\degree}$) will only be accessible for a subset of exoplanets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 4
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics{figures/figure-4-scatterplot.pdf}
    \script{create-figure-4-scatterplot.py}
    \caption{
        Scatter plot for the \HWO target stars, showing stellar effective temperatures and distances.
        %
        The circle sizes represent the angular separation between the star and its planet (in mas) as presented in the target list.
        %
        The colours show $\Delta \alpha$ for hypothetical planets on circular, edge-on orbits at a semi-major axis $a$ corresponding to an Earth-like instellation and an IWA of \qty{62}{\mas}.
        %
        The colour bar indicates the optical phenomena that can in principle be detected: the phenomena from the bottom of the bar up to the colour of a given circle would be detectable for that planet.
        %
        For example, dark blue circles indicate planets for which most phenomena would be observable: planets that can be observed at the rainbow angle can also be observed at angles exhibiting Rayleigh scattering.
    }
    \label{fig:scatterplot}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Circular, inclined orbits}

Most planets are not in edge-on orbits as seen from Earth.
For an orbit with an inclination angle $i$ ($i=0^\circ$ is face-on), we can identify the following two regimes:
\begin{enumerate}
    \item The coronagraph does not obscure any part of the sky-projected orbit: 
    the maximum phase angle coverage depends solely on $i$. 
        This holds for orbits that are relatively face-on (small $i$). 
    \item The coronagraph does obscure parts of the sky-projected orbit: the phase angle coverage depends only on the \IWA and the orbital semi-major axis $a$, following Eq.~\ref{eq:scattering_angle}. 
\end{enumerate}

The range of accessible phase angles is then 
\begin{equation}
\label{eq:Delta_phi_max}
    \Delta \alpha = 
    \begin{cases}
        2 i & \textrm{for} \cos(i) > \frac{\mathrm{IWA}\; d_* }{a} \\ 
        2 \cos^{-1}\left(\dfrac{\mathrm{IWA}\cdot d_* }{a}\right) & \textrm{for} \cos(i) < \frac{\mathrm{IWA}\; d_* }{a}
    \end{cases} \,
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Monte Carlo simulations of accessible orbits}
\label{subsec:2.3}

\subsubsection{Circular, randomly inclined orbits}
\label{sec:circular}

For each of the 164 stars on the \HWO target list, we generate \num{1000} random orbits of hypothetical habitable planets with semi-major axes that yield an Earth-equivalent-incident stellar flux. 
The orbital inclination angles $i$ are randomly drawn from a distribution that is uniform in $\cos i$. 
We then use \cref{eq:Delta_phi_max} to determine the range of accessible phase angles for each hypothetical planet.
The solid lines in \cref{fig:betaallofit} show the cumulative distribution of the maximum and minimum accessible phase angles normalised to the total number of systems in the \HWO target list for different values of the \IWA.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}%[t]
    \centering
    \includegraphics{figures/figure-5-n-over-phase.pdf}
    \script{create-figure-5-n-over-phase.py}
    \caption{
        Cumulative distributions of the most extreme phase angles accessible for different \IWA and for randomly inclined, circular orbits (solid lines) and randomly inclined, eccentric orbits (dashed lines).
        %
        The top and bottom $x$-axes indicate the minimum and maximum accessible phase angles, respectively.
        %
        These angles are symmetric about quadrature (90$^\circ$).
        %
        The $y$-axis on the left indicates the number of planetary systems divided by the number of Monte Carlo samples, and is thus normalised to the number of systems on the target list.
        %
        The $y$-axis on the right indicates the number of HZ Earth-like planets that could be imaged at these phase angles, assuming an occurrence rate of $\eta_\oplus = \qty{24}{\percent}$.
        {\color{red}{the plot says 'elliptical orbits' instead of 'eccentric orbits'}}
    }
    \label{fig:betaallofit}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 7: Yarn ball
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics{figures/figure-7-yarnball.pdf}  
    \script{create-figure-7-yarnball.py}
    \caption{
        Examples of the eccentric orbits generated for the stellar sample. Each plot shows 10 examples of orbits randomly drawn for each star.
        The orbits are scaled to ensure an Earth-like incident flux. 
        %
        The concentric, dashed circles indicate inner working angles of \num{20.6}, \num{41.3}, \num{61.9}, and \qty{82.5}{\mas}.
        %, corresponding to 1, 2, 3, and 4\,$\lambda / D$, respectively (assuming $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$).
        %
        Each system has been scaled such that its HZ has the same radius, hence the variation in the values of the \IWA{}s compared to the orbits.
        %
        It is clear that the \IWA can significantly affect the range of observable phases angles for each orbit.
    }
    \label{fig:ball-o-yarn}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Eccentric, randomly inclined orbits}
\label{sec:eccentric}

We repeat the Monte Carlo simulations from \cref{sec:circular}, assuming the orbital eccentricity follows a beta-distribution with shape parameters $a=0.867$ and $b=3.03$ \citep{2013MNRAS.434L..51K}. 
%
We randomly draw \num{1000} orbits for each star and determine the sky-projected orbits following, for example, \citet{2010exop.book...15M}. 
%
Orbital eccentricities are drawn from the beta distribution, and inclination angles are uniform in $\cos i$. 
%
\Cref{fig:ball-o-yarn} shows a sample of the generated orbits.

We then calculate the minimum and maximum accessible phase angles for \IWA{}s of $\qtylist{21; 42; 63; 84}{\mas}$.
%
The dashed lines in \cref{fig:betaallofit} show the normalised cumulative distribution of the accessible phase angles for each value of the \IWA for eccentric orbits. 
%
We also calculate which target planets would be observable at the phase angles where ocean glint, the rainbow, or the Rayleigh peak occur.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulating contrast curves}
\label{subsec:2.4}

A second parameter that is crucial for detecting angular features like rainbows and ocean glint on Earth-like exoplanets is the contrast, i.e., the planet/star flux ratio. 
%
While not the main focus of this paper, we note that the total and polarized flux contrasts depend on the phase angle, due to a combination of the changing illumination fraction of the observable planetary disk and non-isotropic reflection by the planet.
%
We present some simplified examples in Fig.~\ref{fig:contrasts}. 
%
These calculations are by no means complete or exact, given the many variables that go into the modelling of reflected total and polarized fluxes of Earth-like exoplanets \citep{treesandstam2019,trees2022}.

The NASA Exoplanet Exploration Program’s Mission Star List for \HWO includes for every star an estimated contrast for an Earth-twin in the habitable zone.
%
The contrast is given by \citep[see, e.g.,][]{2023A&A...671A.165M}
\begin{equation}
    C(\alpha) 
    = F_p(\alpha) / F_* 
    = A_\mathrm{G} \cdot \phi(\alpha) \cdot \left( R_p / a(\alpha) \right)^2 \,,
    \label{eq:contrast}
\end{equation}
where $A_\mathrm{G}$ is the planet's geometric albedo, $R_p$ its radius, $\phi(\alpha)$ is the phase function at phase angle $\alpha$ (normalized to \num{1.0} at full phase, thus at $\alpha=\qty{0}{\degree}$), and $a$ is the distance between the planet and its host star.
%
The assumptions that went into the contrast estimations for \HWO are that $a$ is constant (the orbits are circular), the geometric albedo $A_{\rm G}$ is \num{0.2}, and that the phase function $\phi(\alpha)$ can be described by a Lambertian phase function \citep{Russell1916}, which assumes isotropic reflection by the planet. 
%
We assume that the calculated contrasts in the list correspond to the reflected flux at quadrature and use Eq. \ref{eq:contrast} to calculate the reflected light fraction at $\alpha=\qty{90}{\degree}$, i.e. $C(90) = 6.37\%$.

We use the phase functions at $\qty{670}{\nano\meter}$ from \citet{treesandstam2019} for an Earth-like planet with an ocean surface, a wind-speed of $\qty{7}{\meter\per\second}$, and patchy liquid water clouds. 
%
We correct the Star List contrast by dividing by the $6.37\%$ reflected light fraction of the Lambertian phase function and multiply by the reflected light fraction of the \citet{treesandstam2019} model of $5.64\%$.
%
We map the orbital phase for circular orbits to the phase angle and on-sky separation given an inclination $i$ of 90$^\circ$. 
%
Together, they give the reflected light contrast as a function of separation for a full orbit.
%
We multiply this flux with the degree of polarization to obtain the polarized contrast.
%
\cref{fig:contrasts} shows the resulting contrast curves for three of the closest target stars: the M~dwarf Lalande~21185, the K~dwarf $\epsilon$~Eri, and the G~dwarf $\alpha$~Cen~A

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 6
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/figure-6-contrast-over-separation_i=90.pdf}
    \script{create-figure-6-contrast-over-separation.py}
    \caption{
        The planet-to-star contrast ratio and orbital separation for an Earth-like planet with an ocean and patchy clouds along its orbit with an inclination angle $i$ of \qty{90}{\degree} around $\alpha$~Cen~A (G2V), $\epsilon$~Eri (K2V), and Lalande~21185 (M2V).
        %
        The solid, black lines indicate the contrast in unpolarized light, with the contrast at quadrature marked by the black dot.
        %
        The polarized contrast (at $\lambda = \qty{670}{\nano\meter}$) is indicated by the coloured dots, with the colour representing the degrees from quadrature.
        %
        The light grey dots indicate the contrasts at quadrature of the other targets in the star list, and the dashed lines \IWA's of $\qtylist{21; 42; 63}{\mas}$.
    }
    \label{fig:contrasts}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The effect of eccentricity}
\label{sec:result_eccentricity}

As shown in \cref{fig:betaallofit}, allowing for eccentric orbits increases the number of systems that can be observed at each phase angle (notably for the red and purple lines). 
%
This is because eccentric planets spend more time at orbital separations greater than their semi-major axis, and hence are more likely to peek out from behind the coronagraph \IWA and extreme orbital phases (albeit with worse contrast due to the greater star-planet separation).
%
The eccentricity does not significantly affect the number of systems for \IWA smaller than that used in the selection of the \HWO target list (yellow and blue lines).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accessible scattering phenomena}
%\label{sec:results_scattering_phenomena}

\Cref{fig:accessible_phase_angles} and Table \ref{tab:nstars_detect} show the number of systems in which key scattering phenomena will be observable for different \IWA.
%
The Rayleigh peak, which can be used to determine if a planet has an atmosphere, appears near quadrature and is therefore observable in the most systems.
The rainbow feature appears further from quadrature (around $\alpha=\qty{40}{\degree}$), but will still be observable in a significant fraction of systems.
%
The position and shape of the rainbow peak is sensitive to the properties of the cloud droplets, allowing their characterization if the peak of the rainbow feature can be reached.
%
In most systems, only the first half of the glint feature would be observable, but this can still be used to identify a liquid ocean on the surface of a planet, a key indicator of habitability.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
    \centering
    \caption{
        The number of stars around which the peak phase angle of each phenomenon would be detectable, as a function of inner working angle (IWA). %expressed in \unit{\mas} and $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
        %\todo{Do these numbers refer to circular or elliptical orbits?}
        %\todo{Many of these values (e.g., glint) don't really match Fig. 5?}
        {\color{red}{mention 164 total?}}
    }
    \label{tab:nstars_detect}
    % \begin{tabular}{ l c c c } 
    %     \toprule
    %     & \multicolumn{3}{c}{Inner Working Angle}\\
    %     & \qty{21}{\mas} & \qty{41}{\mas} & \qty{62}{\mas} \\
    %     & $2\,\lambda/D$ & $3\,\lambda/D$ & $4\,\lambda/D$\\
    %     \midrule
    %     \midrule
    %     Rainbows    & 90  & 42    & 18 \\
    %     Rayleigh    & 152 & 115   & 65 \\
    %     Ocean Glint & 43  & 14    & 5 \\
    %     % Glory       & 0   & 5     & 10 \\
    %     \bottomrule
    % \end{tabular}\\[5mm]
    %\todo{Just for comparison, here are some values from Fig. 5 (elliptical orbits) that I pulled straight out of the plotting script:}\\[5mm]
    \begin{tabular}{lcccc}
    \toprule
    & \multicolumn{4}{c}{Inner Working Angle (IWA)}\\
    \cmidrule(lr){2-5}
     Feature     &   21 mas &   41 mas &   62 mas &   83 mas \\
    \midrule
     Glory       &        0 &        0 &        0 &        0 \\
     Ocean Glint &       79 &       43 &       16 &        6 \\
     Rainbow     &      109 &       91 &       47 &       23 \\
     Rayleigh Peak   &      154 &      152 &      128 &       82 \\
    \bottomrule
    \end{tabular}\\[5mm]
    %\todo{There is also a mismatch with Fig. 8, though. Where does this discrepancy come frome? Something about phase angle vs. scattering angle?}\\
    %\todo{Observation: The values in the original table seem like a good match for Fig.~8 if we assume the colums where "shifted by one", i.e., 62 mas is actually 83 mas, 41 mas is 62 mas, ...}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contrast curves}
\label{sec:results_contrast}

The contrast of the planet is also an important factor in assessing the detectability of these scattering phenomena. 
%
\Cref{fig:contrasts} shows approximate contrast curves for three of the stars in the \HWO target list assuming an Earth-like planet on an edge-on orbit.
%
The overall contrast varies significantly for each system and throughout the orbit. In general, planets orbiting high-mass stars tend to offer larger angular separation but lower contrast ratios. 
%
The observability of these planets will depend on the starlight suppression of the coronagraph, which is influenced by its \IWA.
%
It will be important to consider both the accessible phase angles and contrast limits when deciding on the \IWA of the \HWO coronagraph, if these scattering features are to be observed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 8
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics{figures/figure-8-n-over-iwa.pdf}
    \script{create-figure-8-n-over-iwa.py}
    \caption{
        The number of systems for which key optical phenomena (Rayleigh peak, rainbow, ocean glint, glory) would be observable as functions of the \IWA.
        %
        Each subplot contains three lines showing the number of systems where the start of each the phenomenon is observable (dashed lines), where the peak is observable (solid lines), and where the end is observable (dotted lines).
        %
        The $x$-axis at the bottom shows the \IWA of the coronagraph in \unit{mas}, and the $x$-axis at the top shows the same \IWA converted to $\lambda/D$ for $\lambda = \qty{600}{\nano\meter}$ and $D = \qty{6}{\meter}$.
        %
        The $y$-axis on the right indicates the number of systems where each phenomena would be observable, assuming \qty{24}{\percent} of them have an Earth-like exoplanet in their HZ.
    }
    \label{fig:accessible_phase_angles}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Habitable Worlds Observatory is a coronagraphic space telescope envisioned to detect and characterize Earth-like exoplanets in the 2040's. 
%
The instrument design and capabilities are still under development and should be driven by the mission's science goals.
%
In this work, we investigated the possibility of using \HWO to detect optical phenomena with specific angular features in polarized and unpolarized reflected light, such as rainbows, ocean glint, and glory, which could establish a planet's habitability. 
%
We primarily focused on the range of accessible phase angles given the limitations of the coronagraphic \IWA and aimed to investigate which phenomena could potentially be observed.
%
While we do not consider the detect-ability of the scattering features beyond their accessibility, we show they are typically more distinct in polarised light but the exoplanet-star contrast ratio is lower than the unpolarised light contrast ratio.
%
A benefit of observing the polarized reflected light is that a coronograph is not explicitly required for quiescent stars as the light of these stars themselves can be considered to be unpolarized \citep{kemp1987}.

The desire to detect and characterize a large number of Earth-like exoplanets, and the intrinsic occurrence of such planets around different stellar spectral types, will drive the choice of coronagraphic \IWA for \HWO. 
%
Assuming a \qty{62}{\mas} \IWA, {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of planetary systems in the NASA Exoplanet Exploration Program's Mission Star List for \HWO would be observable at phases angles where angular features due to Rayleigh scattering show up.
%
The rainbow of light that is scattered in droplets of water clouds and ocean glint, both of which indicate the presence of a water cycle which provides additional context to the habitability of an exoplanet, would be accessible in {\color{red}{!!}} $\sim\qty{-1}{\percent}$ and {\color{red}{!!}} $\sim\qty{-1}{\percent}$ of systems respectively.
%
The presence of a water cycle on Earth-like exoplanets in the remaining systems would have to be established through less direct methods: spectropolarimetry or rotational mapping.
%
To observe more scattering feature in more systems, \HWO' would need to have a smaller \IWA by use of smaller coronograph and potentially a larger primary mirror.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Software/Data Availability}

\todo{check the naming and references/footnotes for this list throughout the paper}
The NASA ExEP target list containing the data used for these simulations is  available online.%
\footnote{\url{https://exoplanets.nasa.gov/exep/science-overview/}}
This work has made use of \textsf{numpy}
 \citep{NumPy2020}, \textsf{scipy} \citep{scipy_2020}, \textsf{matplotlib} \citep{matplotlib2007}, and \textsf{astropy},\footnote{\url{https://www.astropy.org}} a community-developed core Python package and an ecosystem of tools and resources for astronomy \citep{astropy:2013, astropy:2018, astropy:2022}.

This paper has been formatted in the \texttt{showyourwork!} framework \citep{Luger2021}, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible.%
\footnote{\url{https://github.com/mkenworthy/HWObows/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

The 2023 \emph{Optimal Exoplanet Imagers} workshop, which sparked the work presented in this manuscript, was made possible thanks to the logistical and financial support of the Lorentz Center, Leiden, The Netherlands.
%
The research presented in this paper was initiated at a workshop held in Leiden and partially supported by NOVA (the Netherlands Research School for Astronomy) and by the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (grant agreement n°866001 - EXACT).

This research has made use of NASA's Astrophysics Data System Bibliographic Services and the SIMBAD database, operated at CDS, Strasbourg, France.

SRV and JLB acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation program under grant agreement n°805445.
SLC acknowledges support from an STFC Ernest Rutherford Fellowship. 
TDG acknowledges funding from the Max Planck ETH Center for Learning Systems.
KB acknowledges support from NASA Habitable Worlds grant n°80NSSC20K152, and previous support for related work from NASA Astrobiology Institute's Virtual Planetary Laboratory under Cooperative Agreement n°NNA13AA93A.
IL acknowledges the support by a postdoctoral grant issued by the Centre National d'Études Spatiales (CNES) in France.
PB, IL, and YG were supported by the Action Spécifique Haute Résolution Angulaire (ASHRA) of CNRS/INSU co-funded by CNES.
EHP is supported by the NASA Hubble Fellowship grant \#HST-HF2-51467.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
LA and EC acknowledge funding from the European Research Council (ERC) under the European Union's Horizon Europe research and innovation programme (ESCAPE, grant agreement n°101044152).
OA and LK acknowledge funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement n°819155)
SYH was funded by the generous support of the Heising-Simons Foundation.
RJP is supported by NASA under award number 80GSFC21M0002.
OHS acknowledges funding from the direction scientifique générale de l'ONERA (ARE Alioth).

This paper has been formatted in the {\tt showyourwork!} framework, which includes the computer code to reproduce the figures, including the associated data, within an online repository to make it reproducible \citep{Luger2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% References
\bibliographystyle{mnras}
\bibliography{bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Author affiliations
% DO NOT EDIT THIS MANUALLY
% ALL ENTRIES ARE GENERATED BY A SCRIPT
% IF YOU CHANGE THINGS HERE IT WILL GET OVERWRITTEN
\section*{Author affiliations}
\begingroup
    \itshape
    \raggedright
    \affdest{1}Department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK \\[0.5mm]
    \affdest{2}Department of Earth and Planetary Sciences, University of California, Riverside, CA 92521, USA \\[0.5mm]
    \affdest{3}NASA Nexus for Exoplanet System Science, Virtual Planetary Laboratory Team, Seattle, WA 98195, USA \\[0.5mm]
    \affdest{4}NASA Nexus for Exoplanet System Science, Terrestrial Polarization Team, Orlando, FL 32826, USA \\[0.5mm]
    \affdest{5}Centre for Exoplanet Research, School of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK \\[0.5mm]
    \affdest{6}Department of Earth \& Planetary Sciences and Department of Physics, McGill University, 3600 rue University, Montréal, QC, H3A 2T8, Canada \\[0.5mm]
    \affdest{7}Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA Leiden, The Netherlands \\[0.5mm]
    \affdest{8}SRON Netherlands Institute for Space Research, Niels Bohrweg 4, 2333 CA, Leiden, The Netherlands \\[0.5mm]
    \affdest{9}Max Planck Institute for Intelligent Systems, Max-Planck-Ring 4, 72076 Tübingen, Germany \\[0.5mm]
    \affdest{10}ETH Zurich, Institute for Particle Physics \& Astrophysics, Wolfgang-Pauli-Str. 27, 8092 Zurich, Switzerland \\[0.5mm]
    \affdest{11}LESIA, Observatoire de Paris, Université PSL, CNRS, Sorbonne Université, Université de Paris, 92195 Meudon, France \\[0.5mm]
    \affdest{12}Department of Physics, University of California, Santa Barbara, CA 93106, USA \\[0.5mm]
    \affdest{13}STAR Institute, Université de Liège, Allée du six Août 19c, 4000 Liège, Belgium \\[0.5mm]
    \affdest{14}Aix Marseille Univ, CNRS, CNES, LAM, Marseille, France \\[0.5mm]
    \affdest{15}NASA Ames Research Center, Bldg. 245, Moffett Field, USA \\[0.5mm]
    \affdest{16}Université Grenoble Alpes, CNRS, IPAG, 38000 Grenoble, France \\[0.5mm]
    \affdest{17}Department of Astronomy, California Institute of Technology, Pasadena, CA, 91125, USA \\[0.5mm]
    \affdest{18}University of California, Santa Cruz, USA \\[0.5mm]
    \affdest{19}DTIS, ONERA, Université Paris Saclay, 91123 Palaiseau, France \\[0.5mm]
    \affdest{20}DOTA, ONERA, 92322 Châtillon, France \\[0.5mm]
    \affdest{21}Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85719, USA \\[0.5mm]
    \affdest{22}Subaru Telescope, NAOJ, USA \\[0.5mm]
    \affdest{23}College of Optical Sciences, University of Arizona, Tucson, AZ 87521, USA \\[0.5mm]
    \affdest{24}Astrobiology Center, 2 Chome-21-1, Osawa, Mitaka, Tokyo, 181-8588, Japan \\[0.5mm]
    \affdest{25}NASA Goddard Space Flight Center, 8800 Greenbelt Rd, Greenbelt, MD  20771, USA \\[0.5mm]
    \affdest{26}University of Maryland Baltimore County, 1000 Hilltop Cir, Baltimore, MD 21250, USA \\[0.5mm]
    \affdest{27}Faculty of Aerospace Engineering, TU Delft, Building 62, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
    \affdest{28}European Space Agency, ESTEC, Keplerlaan 1, 2200 AG Noordwijk, The Netherlands \\[0.5mm]
    \affdest{29}Université Côte d'Azur, Observatoire de la Côte d'Azur, CNRS, Laboratoire Lagrange, France \\[0.5mm]
    \affdest{30}Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA \\[0.5mm]
    \affdest{31}Delft University of Technology, Kluyverweg 1, 2629 HS Delft, The Netherlands \\[0.5mm]
    \affdest{32}Department of Geoscience \& Remote Sensing, Delft University of Technology, Stevinweg 1, 2628 CN, Delft, The Netherlands \\[0.5mm]
    \affdest{33}Royal Netherlands Meteorological Institute (KNMI), Utrechtseweg 297, 3731 GA, de Bilt, The Netherlands \\[0.5mm]
    \affdest{34}National Research Council Canada, Herzberg Astronomy and Astrophysics Research Center, Victoria, B.C., Canada \\[0.5mm]
\endgroup

\end{document}